---
title: "Streetlight"
output: html_notebook
---

## Setup

```{r setup}

knitr::opts_knit$set(root.dir = "C:/Users/dgallen/Desktop/R/Shiny/Streetlight/TH_36_Manning_Expanded_Updated_8041_Travel")

library(tidyverse)
library(lubridate)
library(gridExtra)
library(reshape2)
library(zoo)
library(rgdal)
library(readtext)
library(leaflet)

```

Look at the files in the folder:

```{r files in folder}

dir()

```

We have:

* 2 .txt files. From inspecting directly, one is a description of the variables and one includes details of the data pull
* 5 .csv files containing the traffic data.
* 3 shape files showing the origin destination and pass through zones.

## Text File

We will organise the "Project" text file. It contains information regarding the data pull from the Streetlight website and the data itself.

```{r Project.txt,warning=FALSE}

project <- read.table("Project_OD_MF.txt",sep='\n',stringsAsFactors = FALSE)%>%
  separate(V1,c('variable','value'),':')

project$Info <- ifelse(unlist(lapply(project$variable,function(x) any(x==c(0:10))))==TRUE,NA,project$variable)%>%
  na.locf()

project_details <- select(project,Info,value)%>%
  filter(value!=' ')
rm(project)
print(project_details)
```

We my need this data later, so now we have it in a format we can use. We may need to dcast and remelt to fix the day types and day types.

## Data files 

### mf_commerical

We examine the first csv file. We want to set a function that will find a the right file regardless of the name. 

```{r mf_commerical read data}

mf_com_names <- dir()[grepl("*_mf_commercial.csv",dir())]

mf_com_dat <- read.csv(mf_com_names[!grepl('zone',mf_com_names)],stringsAsFactors = FALSE)

mf_com_zone <- read.csv(mf_com_names[grepl('zone',mf_com_names)],stringsAsFactors = FALSE)

str(mf_com_dat)

```

Lets sgtart at the end with the most interesting field; "Avg.Trip.Duration..sec". We want to check the data for outliers, missing values or interesting trents

```{r mf_com_dat avg trip duration}

### change variable to numeric


mf_com_dat$Avg.Trip.Duration..sec. <-  as.numeric(mf_com_dat$Avg.Trip.Duration..sec.)

hist(mf_com_dat$Avg.Trip.Duration..sec.,breaks = 200)

table(is.na(mf_com_dat$Avg.Trip.Duration..sec.))

```

Looks like we have some extreme outliers and 10% NAs which is significant.

Lets invesitgate the NAs.

```{r mf_com_dat NAs}


head(mf_com_dat[is.na(mf_com_dat$Avg.Trip.Duration..sec.),])
tail(mf_com_dat[is.na(mf_com_dat$Avg.Trip.Duration..sec.),])
```

Unclear as to why these values are NA. Check with the experts. 

We now melt down the commerical dataset. We want each row to be an origin, a pass_through or a destination. We can now easily use group by operations. Wide to long.

```{r melt down mf coms}

mf_com_dat_melt <- melt(mf_com_dat,measure.vars = c("Origin.Zone.ID","Origin.Zone.Name","Middle.Filter.Zone.ID","Middle.Filter.Zone.Name","Destination.Zone.ID","Destination.Zone.Name"),id.vars = c("Device.Type","Day.Type","Day.Part","O.M.D.Traffic..StL.Index.","Origin.Zone.Traffic..StL.Index.","Middle.Filter.Zone.Traffic..StL.Index.","Destination.Zone.Traffic..StL.Index.","Avg.Trip.Duration..sec."))

mf_com_dat_melt$group <- ifelse(grepl("Origin",mf_com_dat_melt$variable),"From",
                                ifelse(grepl("Middle",mf_com_dat_melt$variable),"Through",
                                       ifelse(grepl("Destination",mf_com_dat_melt$variable),"To",NA)))
```

Each row is now as we want.

Create an OD matrix

```{r}
mf_com_dat_OD <- select(mf_com_dat,c("Origin.Zone.ID","Origin.Zone.Name","Middle.Filter.Zone.ID","Middle.Filter.Zone.Name","Destination.Zone.ID","Destination.Zone.Name","Device.Type","Day.Type","Day.Part","O.M.D.Traffic..StL.Index.","Origin.Zone.Traffic..StL.Index.","Middle.Filter.Zone.Traffic..StL.Index.","Destination.Zone.Traffic..StL.Index.","Avg.Trip.Duration..sec."))


total <-  sum(mf_com_dat_OD$Avg.Trip.Duration..sec.)

mf_com_dat_OD_matrix <-  mf_com_dat_OD%>%
  filter(Day.Type=='0: Average Day (M-Su)',
         Day.Part=='0: All Day (12am-12am)',
         Middle.Filter.Zone.Name=='TH 36 WB Stillwater Blvd On-Ramp')%>%
  select(Origin.Zone.Name,Destination.Zone.Name,Avg.Trip.Duration..sec.)%>%
  mutate(Avg.Dur.Perc. = round(Avg.Trip.Duration..sec./(sum(Avg.Trip.Duration..sec.,na.rm = TRUE))*100,2))%>%
  select(Destination.Zone.Name,Avg.Dur.Perc.,Origin.Zone.Name)%>%
  spread(key=Destination.Zone.Name,value=Avg.Dur.Perc.)%>%
  mutate(Total.Perc = rowSums(.[-1],na.rm = T))

```


### mf_personnal


```{r mf_personal read data}

mf_pers_name <- dir()[grepl("*_mf_personal.csv",dir())][1]

mf_pers <- read.csv(mf_pers_name,stringsAsFactors = FALSE)
dir()
str(mf_pers)

```

Lets start at the end with the most interesting field; "Avg.Trip.Duration..sec". We want to check the data for outliers, missing values or interesting trents

```{r mf_pers avg trip duration}

### change variable to numeric


mf_pers$Avg.Trip.Duration..sec. <-  as.numeric(mf_pers$Avg.Trip.Duration..sec.)

hist(mf_pers$Avg.Trip.Duration..sec.,breaks = 200)

table(is.na(mf_pers$Avg.Trip.Duration..sec.))

```

```{r melt down mf pers}

mf_pers_melt <- melt(mf_pers,measure.vars = colnames(mf_pers)[2:15],id.vars = colnames(mf_pers)[c(1,16:22)])

mf_pers_melt$group <- ifelse(grepl("Origin",mf_pers_melt$variable),"From",
                            ifelse(grepl("Middle",mf_pers_melt$variable),"Through",
                            ifelse(grepl("Destination",mf_pers_melt$variable),"To",NA)))
```


### Combined Data

We combine the commerical and personnal dataset. 

```{r bind pers and com}

mf <- rbind(mf_pers_melt,mf_com_dat_melt)

```



## Shapefiles

Load the shapefiles

There are 3 shapefiles in the directory, again one for the origin, middle and destination.

### Origin Shapefile

```{r shape origin}
shape_org_name <- dir()[grepl("*_origin_zone_set.cpg*",dir())]
shape_org_name <- substr(shape_org_name,1,nchar(shape_org_name)-4)

shape_org <- rgdal::readOGR(dsn='.',layer = shape_org_name)
```

```{r}
trueCentroids <- SpatialPointsDataFrame(rgeos::gCentroid(shape_org_merge,byid = TRUE),shape_org_merge@data,match.ID = FALSE)%>%as.data.frame()
```


```{r leaflet origin}

icons <- awesomeIcons(
  icon = 'ion-model-s',
  iconColor = 'red',
  library = 'ion'
)

total_org <- mf_com_dat_OD%>%
  filter(Day.Type=='0: Average Day (M-Su)',
         Day.Part=='0: All Day (12am-12am)',
         Middle.Filter.Zone.Name=='TH 36 WB Stillwater Blvd On-Ramp')%>%
  select(Origin.Zone.ID,Destination.Zone.Name,Avg.Trip.Duration..sec.)%>%
  mutate(Avg.Dur.Perc. = round(Avg.Trip.Duration..sec./(sum(Avg.Trip.Duration..sec.,na.rm = TRUE))*100,2))%>%
  select(Destination.Zone.Name,Avg.Dur.Perc.,Origin.Zone.ID)%>%
  spread(key=Destination.Zone.Name,value=Avg.Dur.Perc.)%>%
  mutate(Total.Perc = rowSums(.[-1],na.rm = T))%>%
  select(Origin.Zone.ID,Total.Perc)


labels_org <- paste('ID:',shape_org$idsha,', ', shape_org$name,sep='')

shape_org_merge <- merge(shape_org,total_org,by.x='id',by.y='Origin.Zone.ID',all.x=TRUE)

leaflet(shape_org) %>%
  addTiles(group = "OSM (default)") %>%
  addPolygons(color = "red", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    highlightOptions = highlightOptions(color = "white", weight = 5,
      bringToFront = TRUE),label=labels_org)%>%
  addCircleMarkers(lat=~trueCentroids$y,lng=~trueCentroids$x
                   ,radius=~trueCentroids$Total.Perc)
```



### Destination Shapefile

```{r shape destination}
shape_dest_name <- dir()[grepl("*_destination_zone_set.cpg*",dir())]
shape_dest_name <- substr(shape_dest_name,1,nchar(shape_dest_name)-4)

shape_dest <- rgdal::readOGR(dsn='.',layer = shape_dest_name)

```

```{r leaflet destination}


labels_dest <- paste('ID:',shape_dest$id,', ', shape_dest$name,sep='')

leaflet(shape_dest) %>%
  addTiles(group = "OSM (default)") %>%
  addPolygons(color = "red", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    highlightOptions = highlightOptions(color = "white", weight = 5,
      bringToFront = TRUE),label=labels_dest)

```






### Middle Shapefile

```{r shape middle}
shape_mid_name <- dir()[grepl("*_filter_zone_set.cpg*",dir())]
shape_mid_name <- substr(shape_mid_name,1,nchar(shape_mid_name)-4)

shape_mid <- rgdal::readOGR(dsn='.',layer = shape_mid_name)

```

```{r leaflet middle}

labels_mid <- paste('ID:',shape_mid$id,', ', shape_mid$name,sep='')

leaflet(shape_mid) %>%
  addTiles(group = "OSM (default)") %>%
  addPolygons(color = "red", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    highlightOptions = highlightOptions(color = "white", weight = 5,
      bringToFront = TRUE),label=labels_mid)
```

## Circlize



